I/O simulation model with Linux page cache, integration and evaluation in SimGrid framework

### 2020-10-12 paper_io_sim.tex
Export svg figures
Write results of experiment 1, 2

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{hyperref}

@@

			We use our dedicated cluster hosted at Concordia University to conduct 
			the experiments. The cluster consists of one login node, 8 compute nodes 
			and 4 storage nodes connected with two network switches. The login node 
			has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 137~GB (128~GiB) of RAM, 
			1.8TB of storage of XFS file system, 13~GB of tmpfs file system and 385 TB of 
			Lustre shared file system. Each compute node has two 16 cores Intel(R) 
			Xeon(R) Gold 6130 CPU @ 2.10GHz, 275~GB (256~GiB) of RAM and 6 SSDs, 
			450~GB each with XFS file system, 378~GB of tmpfs and 126~GB of devtmpfs file system.
			Lustre file system is configured with one metadata target with 854~GB 
			of storage, 44 object storage targets with 8.7~TB of storage each. 
			The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed. 
			We use \textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to 
			monitor and collect data of memory, page cache status and disk throughput 
			on the cluster. 
			\textcolor{red}{[Add network description here, include number of disks 
			on all nodes ,maybe add a diagram]}

@@

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{caption}
\usepackage{subcaption}

@@

			We use our dedicated cluster hosted at Concordia University to conduct 
			the experiments. The cluster consists of one login node, 8 compute nodes 
			and 4 storage nodes connected with two network switches. The login node 
			has one Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz, 137~GB (128~GiB) of RAM, 
			1.8TB of storage of XFS file system, 13~GB of tmpfs file system and 385 TB of 
			Lustre shared file system. Each compute node has two 16 cores Intel(R) 
			Xeon(R) Gold 6130 CPU @ 2.10GHz, 275~GB (256~GiB) of RAM and 6 SSDs, 
			450~GB each with XFS file system, 378~GB of tmpfs and 126~GB of devtmpfs file system.
			Lustre file system is configured with one metadata target with 854~GB 
			of storage, 44 object storage targets with 8.7~TB of storage each. 
			The cluster is run on CentOS 8.1 with the Slurm Workload Manager installed. 
			We use \textbf{\textit{atop}} and \textbf{\textit{collectl}} as tools to 
			monitor and collect memory status and disk throughput 
			on the cluster. 
			\textcolor{red}{[Add network description here, include number of disks 
			on all nodes ,maybe add a diagram]}

@@

		We use the disk and memory benchmark results and CPU time measured from 
		the cluster to configure the simulated platform. The benchmark results of 
		bandwidths and configured values in the simulators are shown in \ref{table:benchmark}.

		\begin{table}
		\centering
		\begin{tabular}{|l|c|c|c|}
		\hline
			Bandwidths (MBps)  & Cluster & Python & WRENCH\\
		\hline
			Memory read  & 6860	& 6860	 & 2765\\
			Memory write & 2765	& 2765 & 2765\\
			Disk read & ~510 & 465 & 465\\
			Disk write & ~420 & 465	 & 465\\
		\hline
		\end{tabular}
		\caption{Bandwidth benchmark results and simulator configurations}
		\label{table:benchmark}
		\end{table}

		In the Python simulator, we use different read and write bandwidths for both 
		memory and disk as in cluster. However, as read and write bandwidths are 
		symmetrical in SimGrid, we configure memory with write bandwidth 
		and disk with average bandwidth.

		\subsection{Single-threaded experiment}

		\subsubsection{Simulated time}

			\begin{figure*}
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/error_20.pdf}
   				\caption{20GB input file}	\label{fig:single_error_20}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/error_50.pdf}
   				\caption{50GB input file}	\label{fig:single_error_50}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/error_75.pdf}
   				\caption{75GB input file}	\label{fig:single_error_75}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/error_100.pdf}
   				\caption{100GB input file}	\label{fig:single_error_100}
			\end{subfigure}	
			\caption{Relative error of individual simulated tasks}
			\label{fig:single_error}
			\end{figure*}

			The I/O time of tasks in both simulators is tracked, and then compared 
			to the I/O time in the real pipeline execution. 
			Figure~\ref{fig:single_error} shows the relative error of simulated I/O time of 
			individual I/O activities in simulators with different input sizes. 
			There are 3 tasks in the pipeline and each task reads an input file, 
			performs some computation, and writes output file. 
			Thus, we have 6 sequential I/O activities, \textit{read1} and \textit{write1} 
			from the first task, \textit{read2} and \textit{write2} the second task, 
			and \textit{read3}, \textit{write3} from the last task.

			In general, with all input sizes used in the experiment, the simulation errors 
			of our simulators, both Python and WRENCH-Ext are remarkably lower 
			in comparison with that of the original WRENCH simulator . 
			The reason for this is, in original WRENCH, because page cache is not simulated, 
			all data is only read/written with disk bandwidth, which makes simulated 
			I/O time longer than in reality.
			Especially, in our simulators, the errors of the Python simulator are significantly 
			higher than that of the WRENCH-Ext simulator. 
			This difference may attribute to the limitation in the periodical flushing 
			mechanism simulation in Python. 
			In WRENCH, it can be easily implemented with a simulated background thread, 
			while there is only one thread simulated in our Python simulator, which 
			hinders the correctness of this mechanism. 
			Noticeably, the simulation time errors of \textit{read3} in the Python and 
			WRENCH-Ext simulators are just under the error in original WRENCH. 
			We will give our explanation for this later in this section.

		\subsubsection{Memory behavior}

			\begin{figure*}
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_20gb.pdf}
   				\caption{20GB input file}	\label{fig:single_memprof_20gb}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_50gb.pdf}
   				\caption{50GB input file}	\label{fig:single_memprof_50gb}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_75gb.pdf}
   				\caption{75GB input file}	\label{fig:single_memprof_75gb}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/single_comp_sep_100gb.pdf}
   				\caption{100GB input file}	\label{fig:single_memprof_100gb}
			\end{subfigure}	
			\caption{Memory profiling results with different input file sizes}
			\label{fig:single_memprof}
			\end{figure*}

			Memory behaviors of real pipeline execution and simulators are also recorded 
			and the memory profiling results are illustrated in Figure~\ref{fig:single_memprof}, 
			with the bars presenting activities in tasks (read, compute and write) 
			and the lines describing memory status throughout pipeline run time.

			As shown in the figures, the makespans of the simulated pipeline are 
			close to the makespan of the real pipeline. 
			But more importantly, with all input file sizes, the trends of the lines describing 
			memory in simulators are very similar to observations in the real execution. 
			The similar trends in the amount of cache used (purple lines) in both reality 
			and simulation shows that the use of page cache is well reproduced in the simulators. 
			Besides, the red lines, which represent dirty data in cache, also 
			follow similar trends. That means not only the amount dirty data written to 
			cache is simulated, but also the flushing and periodical flushing 
			mechanisms are accurately modeled in our simulators. 
			This can be seen in the decrease in the amount of dirty data during computation 
			time as well as read time. 
			In addition, the effect of \textit{dirty\_data}  threshold is also simulated, 
			which is reflected the interactions between dirty data (red line) and dirty\_ratio 
			(black dashed line).

			\begin{figure*}
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_real.pdf}
   				\caption{Real pipeline execution}	
   				\label{fig:fincore_real}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/single/figures/fincore_wrench.pdf}
   				\caption{WRENCH Simulator}	
   				\label{fig:fincore_wrench}
			\end{subfigure}	
			\caption{Amount of file data in cache after each I/O activity with 100GB of input}
			\label{fig:fincore}
			\end{figure*}

			Again, with 100GB input file, memory behavior in the third read in the 
			simulators is different from reality, causing unexpected long read time, 
			which we have seen in Figure~\ref{fig:single_error_100}. 
			This can be explained with the amount of cached data of files at the beginning 
			of the read shown in Figure~\ref{fig:fincore}. 
			In real pipeline execution, after \textit{write2}, which writes \textit{file3}, 
			all data of \textit{file3} is in cache (Figure~\ref{fig:fincore_real}). 
			Thus, in \textit{read3}, \textit{file3} is read with memory bandwidth. 
			But in our simulator, after \textit{write2}, only a part of \textit{file3} 
			remains in cache (Figure~\ref{fig:fincore_wrench}). 
			The reason is while \textit{file3}  is being written, the page cache is saturated, 
			and eviction is evoked to evict data from page cache, make space 
			available for \textit{file3} data. 
			However, because the amount of evictable memory is less than required amount, 
			some \textit{file3} data is evicted after being written, make \textit{file3} 
			partially cached. As a result, in \textit{read3}, only a part of \textit{file3} is read 
			from cache with memory bandwidth, a part of it is read from disk with disk read 
			bandwidth, which is many times slower, leading to longer read time.

			In summary, in single-threaded applications, our model can accurately 
			simulate I/O time as well as the internal memory behavior. 
			The model not only proves its applicability, but also proves that it is on 
			a right track and can be further improve to achieve higher accuracy.

		\subsection{Multi-threaded experiment}

			\begin{figure*}
			\begin{subfigure}{\linewidth}
				\centering
   				\includegraphics[width=0.5\columnwidth]{result/multi/figures/makespan.pdf}
   				\caption{Total makespan}	
   				\label{fig:multi_makespan}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/multi/figures/readtime.pdf}
   				\caption{Cumulative read time}	
   				\label{fig:multi_readtime}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/multi/figures/readtime_avg.pdf}
   				\caption{Average read time}	
   				\label{fig:multi_readtime_avg}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/multi/figures/writetime.pdf}
   				\caption{Cumulative write time}	
   				\label{fig:multi_writetime}
			\end{subfigure}	
			\begin{subfigure}{\columnwidth}
				\centering
   				\includegraphics[width=\columnwidth]{result/multi/figures/writetime_avg.pdf}
   				\caption{Average write time}	
   				\label{fig:multi_writetime_avg}
			\end{subfigure}	
			\caption{Total makespan and I/O time with different number of concurrent pipeline}
			\label{fig:multi_result}
			\end{figure*}

			In this experiment, we vary the number of 	concurrent pipeline from 1 to 32 
			and record total makespan, cumulative and average read/write time 
			of the pipelines on the cluster as well as from original WRENCH and 
			WRENCH-Ext simulators. 
			The results are presented in Figure~\ref{fig:multi_result}.

			The Figure~\ref{fig:multi_makespan} shows that as the number of pipelines 
			increases, the total makespan raises accordingly because the amount of data 
			increases but the memory and disk bandwidths are constant and shared between 
			pipelines. 
			In comparison with original WRENCH, the result from WRENCH-Ext are closer to 
			reality because it is implemented with our I/O with page cache model.

			If we focus on read time only, both cumulative read time (Figure~\ref{fig:multi_readtime}) 
			and average read time (Figure~\ref{fig:multi_readtime_avg}) 
			follow the same trend in total makespan, and the simulated read time 
			of WRENCH-Ext simulator is closer to reality than the read time in original WRENCH. 
			In addition, the average read time seems to increase linearly in all cases when 
			the bandwidth is shared by pipelines. Again, the result from WRENCH-Ext is more 
			accurate than that from original WRENCH.

			Similarly, the cumulative write time (shown in Figure~\ref{fig:multi_writetime}) 
			and the average write time (shown in Figure~\ref{fig:multi_writetime_avg}) have 
			the same increasing trend and the result from WRENCH-Ext are more realistic than 
			the result from original WRENCH. 
			Especially, the average write time in both reality and WRENCH-Ext have a similar 
			pattern, when they gradually increase before a surge linearly with the same slope 
			after the number of pipelines reaches 10. 
			This can be explained that with less than 10 concurrent pipelines, the page cache 
			is still not saturated, so all files can be written entirely to cache with memory 
			bandwidth in a short time. 
			After the page cache is saturated with dirty data (after around 10 concurrent pipelines 
			in this experiment), this dirty data needs to be flushed in order to make space 
			available for new data being written to cache. This dirty data is flushed to 
			disk with disk bandwidth, which is many times slower than memory bandwidth. 
			The more pipelines we have, the more data needs to be written, the more data 
			needs to be flushed, leading to a sharper increase in the average write time.

			To conclude, the results from WRENCH-Ext shows that our model can simulate I/O 
			with page cache accurately in a multi-threaded application.

		\subsection{Remote storage}

		\subsection{Real application}
